<small> Supplementary materials for the main submission entitled - <br><br> A new metric for automatic discovery of periodic patterns in time series

### 2021-05-04 paper/paper-supplementary.Rmd
supplementary section tuning paramter sim study update

# Simulation design

@@

Following the work of Krzysztofowicz (1997) the empirical NQT involves the following steps:
1. Sort the sample of measured variable $X$ from the smallest to the largest observation $x_{(1)},\dots, x_{(i)}, .., x_{(n)}$.
2. Estimate the cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$.
using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(X \leq x_{(i)})$.
3. Transforming each observation $x_(i)$ of $X$ into observation $y(i) = Q^{-1}(p(i))$ of the standard normal variate $Y$ , with $Q$ denoting the standard normal distribution and
$Q^{-1}$ its inverse, applying discrete mapping.

Firstly, we show that NQT works to make the distribution within each panel same. Next, we study the distribution of $wpd$ for different $nx$ and $nfacet$ in various ways.

## Normal quantile transform

Since the measure $wpd$ is essentially set up to detect "differences" in distributions irrespective of underlying distribution, it would be ideal if it has minimal dependency on the type, location and scale of the initial distribution. To that end, some data pre-processing through the Normal Score Transform (NQT) has been applied in order to make most asymmetrical distributed measured variables more normal-like. Figure \ref{fig:gamma-diff-wotrans} shows ridge plots of raw $wpd$ for a Gamma(0.5,1), Gamma(2,1) before NQT. It is observed that for the underlying distribution Gamma(2,1), location and scale of the distribution of wpd changes from top-left panel to bottom-right panel. Moreover, the location and scale of the distribution of $wpd$ fir different underlying distribution Gamma(0.5,1), Gamma(2,1). Figure \ref{fig:gamma-diff-quantrans} shows the the distributions of $wpd$ under same underlying distributions but after performing NQT. It is observed that within each panel, the distributions of the $wpd$ looks same, however, the distributions change from extreme top-left panel to bottom-right panels. This implies, NQT has atleast been able to bridge the gap in distribution of $wpd$ for different non-normal underlying distributions.

@@

## By distribution

```{r normal-diff-quantrans, fig.cap = "Ridge plots of raw wpd is shown for N(0,1), N(5,1) and N(0,5) distribution. The densities change across different facet and x levels but look same for each panel, which implies wpd value is unaffected by the change in mean and standard deviation of the normal distribution", eval = FALSE}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_normal.png"))
```

## By category levels

@@

```{r tuning-oneomega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```

```{r tuning-parameter-inter, fig.cap ="For most panels it is observed that the most common value of the tuning paramter for which the designs interact is 0.5, which implies any value greater than 0.5 could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances for most sitatuions."}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

>>>>

How does the tuning parameter affects the value of $wpd$ under different designs? The tuning parameter is used to put relative weight-age to the difference in distributions within and between facets. So it is interesting to see how the value of $wpd$ changes for these two designs.

### Simulation design

Observations are generated from Normal(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20\}$. 
Let $\omega = \{1, 2, \dots, 10\}$ denotes the variable denoting the increment in mean of the distribution. The values of lambda ranges from ${0.1, 0.2, \dots, 0.9}$. Two designs are considered $D_{var_x}$ and $D_{var_f}$ and the values of $wpd$ is being computed for all these different values of the considered variable.

### Results

Figure \ref{fig:tuning-oneomega} shows how the value of $wpd$ changes for $\lambda = {0.1, 0.2, \dots, 0.9}$ for the two different designs $D_{var_x}$ and $D_{var_f}$ for two values of increment in mean $\omega = {1, 8}$. For a lower value of $\omega$, the two designs intersect at $\lambda > 0.7$ and for a higher $\omega$, the two designs intersect at $\lambda =  0.5$.
The value of $wpd$ increases with $\lambda$ for $D_{var_x}$ and decreases with increasing $\lambda$ for 
$D_{var_f}$. Figure \ref{fig:tuning-parameter-inter} shows the value of $\lambda$ for which the two designs intersect across different values of $\omega$. It can be observed that as the value of $\omega$ ($\omega > 4$) increase, the value of $\lambda$ at which the two designs intersect converge is $\lambda = 0.5$.

```{r tuning-oneomega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```
```{r tuning-parameter-inter, fig.cap ="For most panels it is observed that the most common value of the tuning paramter for which the designs interact is 0.5, which implies any value greater than 0.5 could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances for most sitatuions."}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

# Normalization

## Simulation design

@@

## Different underlying distributions

Since the measure $wpd$ is essentially set up to detect "differences" in distributions irrespective of underlying distribution, it would be ideal if it has minimal dependency on the type, location and scale of the initial distribution. To that end, some data pre-processing through the Normal Score Transform (NQT) has been applied in order to make most asymmetrical distributed measured variables more normal-like. Figure \ref{fig:gamma-diff-wotrans} shows ridge plots of raw $wpd$ for a Gamma(0.5,1), Gamma(2,1) before NQT. It is observed that for the underlying distribution Gamma(2,1), location and scale of the distribution of wpd changes from top-left panel to bottom-right panel. Moreover, the location and scale of the distribution of $wpd$ for different underlying distribution Gamma(0.5,1), Gamma(2,1). Figure \ref{fig:gamma-diff-quantrans} shows the the distributions of $wpd$ under same underlying distributions but after performing NQT. It is observed that within each panel, the distributions of the $wpd$ looks same, however, the distributions change from extreme top-left panel to bottom-right panels. Similar observations could be made in Figure \ref{fig:normal-diff-quantrans} for different underlying normal distributions  N(0,1), N(5,1) and N(0,5). This implies, NQT has atleast been able to bridge the gap in distribution of $wpd$ for different non-normal underlying distributions.

@@

```{r normal-diff-quantrans, fig.cap = "Ridge plots of raw wpd is shown for N(0,1), N(5,1) and N(0,5) distribution. The densities change across different facet and x levels but look same for each panel, which implies wpd value is unaffected by the change in mean and standard deviation of the normal distribution", eval = FALSE}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_normal.png"))
```
## Different number of comparisons

# Location and scale

Next, we study the distribution of $wpd$ for different $nx$ and $nfacet$ in various ways.

@@

<!-- Following the work of Krzysztofowicz (1997) the empirical NQT involves the following steps: -->
<!-- 1. Sort the sample of measured variable $X$ from the smallest to the largest observation $x_{(1)},\dots, x_{(i)}, .., x_{(n)}$. -->
<!-- 2. Estimate the cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$. -->
<!-- using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(X \leq x_{(i)})$. -->
<!-- 3. Transforming each observation $x_(i)$ of $X$ into observation $y(i) = Q^{-1}(p(i))$ of the standard normal variate $Y$ , with $Q$ denoting the standard normal distribution and -->
<!-- $Q^{-1}$ its inverse, applying discrete mapping. -->

<!-- Firstly, we show that NQT works to make the distribution within each panel same. . -->
